Analysis of cloud enhancement events in a 30-year record of global solar irradiance at Thessaloniki, Greece

### 2024-10-07 article/article.Rmd
Commit 2024-10-07 09:12

The cloud-free reference values discussed above are based on the climatological AOD
and WC; hence they cannot describe accurately the long-term variation of
$\text{GHI}_\text{ref}$ due to long-term changes in the two atmospheric constituents,
mainly AOD.  As reported by @Natsis2023, there is a long-term brightening effect in
the GHI data of Thessaloniki for the period 1993 -- 2023, which for cloud-free data
was attributed to long-term changes in aerosol effects.  Therefore, an adjustment of
the $\text{GHI}_\text{ref}$ during the period of study was made using RTM simulations
to account for the long-term variations of the AOD.  High quality AOD data with a
Cimel sun-photometer at Thessaloniki start only in 2003, while spectral AOD
measurements using direct-sun observations with a MKIII Brewer spectrophotometer are
available for the period 1997 -- 2017. The Brewer AOD data are taken sporadically
during each day and are less dense compared to the AERONET data. By comparing monthly
AOD data at $340\,\text{nm}$ of the two instruments for the common periods of
operation we adjusted the Brewer data and filled the missing months of the Cimel
record with Brewer data.  Using the extended monthly time series of AOD at
$340\,\text{nm}$, as well as monthly climatological values of the Ångström exponent
and constant WC of $15.6\,\text{mm}$ derived from the Cimel record, we simulated with
the RTM the cloud-free GHI at SZA of $55^\circ$ for each month in the period 1997 --
2023. The SZA of $55^\circ$ was chosen as representative of all days in the year to
get a rough estimate of the annually averaged change in cloud-free GHI. A
second-degree polynomial fit was applied to the simulated yearly averaged GHI to
derive the long-term change in GHI due to aerosols:
\begin{equation}
\Delta(\text{GHI}) [\%] = -12170 + 12.05 \cdot x + -0.002981 \cdot x^2 \label{eq:AODchange}
\end{equation}
where $x$ is the date as a decimal fraction of the year.  The resulting change in
$\text{GHI}_\text{ref}$ ranges between $-1.88$ in 1994 and $-0.19\,\%$ in the end of
2023, peaking at $0.23\,\%$ in the mid-2020. Finally, we applied these long-term
changes of Equation\nobreakspace{}\ref{eq:AODchange} to the $\text{GHI}_\text{ref}$,
to create a more realistic representation of the cloud-free irradiance for the entire
period of study. For the period 1994 -- 1996 where no AOD data are available, we
assumed that the changes in GHI follow the same polynomial fit.

@@

One of the goals of this study was to quantify the OI related to CEs. A key issue for
achieving this goal is to define a threshold for the CE identification,
representative of the cloud-free irradiance at the time of each GHI measurement. This
depends on the selection of the appropriate atmospheric parameterization for the RTM
simulations. The implementation of the long-term changes in GHI due to AOD, discussed
in section\nobreakspace{}\ref{rtmcs}, allows capturing a large part of the natural
variability of cloud-free GHI. However, the short-term variability of AOD cannot be
taken adequately into account when using monthly values in the model simulations. We
tried different approaches to strengthen the robustness of the methodology and to
compensate for the limited accuracy of the RTM input data and the unpredictable
natural variability of the atmosphere.

>>>>

The cloud-free reference values discussed above are based on the climatological AOD
and WC; hence they cannot describe accurately the long-term variation of
$\text{GHI}_\text{ref}$ due to long-term changes in the two atmospheric constituents,
mainly AOD.  As reported by @Natsis2023, there is a long-term brightening effect in
the GHI data of Thessaloniki for the period 1993 -- 2023, which for cloud-free data
was attributed to long-term changes in aerosol effects.  Therefore, an adjustment of
the $\text{GHI}_\text{ref}$ during the period of study was made using RTM simulations
to account for the long-term variations of the AOD.  High quality AOD data with a
Cimel sun-photometer at Thessaloniki start only in 2003, while spectral AOD
measurements using direct-sun observations with a MKIII Brewer spectrophotometer are
available for the period 1997 -- 2017. The Brewer AOD data are taken sporadically
during each day and are less dense compared to the AERONET data. By comparing monthly
AOD data at $340\,\text{nm}$ of the two instruments for the common periods of
operation we adjusted the Brewer data and filled the missing months of the Cimel
record with Brewer data.  Using the extended monthly time series of AOD at
$340\,\text{nm}$, as well as monthly climatological values of the Ångström exponent
and constant WC of $15.6\,\text{mm}$ derived from the Cimel record, we simulated with
the RTM the cloud-free GHI at SZA of $55^\circ$ for each month in the period 1997 --
2023. The SZA of $55^\circ$ was chosen as representative of all days in the year to
get a rough estimate of the annually averaged change in cloud-free GHI. A
second-degree polynomial fit was applied to the simulated yearly averaged GHI to
derive the long-term change in GHI due to aerosols:
\begin{equation}
\Delta(\text{GHI}) [\%] = -12170 + 12.05 \cdot x -0.002981 \cdot x^2 \label{eq:AODchange}
\end{equation}
where $x$ is the date as a decimal fraction of the year.  The resulting change in
$\text{GHI}_\text{ref}$ ranges between $-1.88\,\%$ in 1994 and $-0.19\,\%$ in the end of
2023, peaking at $0.23\,\%$ in the mid-2020. Finally, we applied these long-term
changes of Equation\nobreakspace{}\ref{eq:AODchange} to the $\text{GHI}_\text{ref}$,
to create a more realistic representation of the cloud-free irradiance for the entire
period of study. For the period 1994 -- 1996 where no AOD data are available, we
assumed that the changes in GHI follow the same polynomial fit.

@@

One of the goals of this study was to quantify the OI related to CEs. A key issue for
achieving this goal is to define a threshold for the CE identification,
representative of the cloud-free irradiance at the time of each GHI measurement.
This depends on the selection of the appropriate values of the input parameters for
the RTM simulations.  The implementation of the long-term changes in GHI due to AOD,
discussed in section\nobreakspace{}\ref{rtmcs}, allows capturing a large part of the
natural variability of cloud-free GHI. However, the short-term variability of AOD
cannot be taken adequately into account when using monthly values in the model
simulations. We tried different approaches to strengthen the robustness of the
methodology and to compensate for the limited accuracy of the RTM input data and the
unpredictable natural variability of the atmosphere.
